---
title: "p8105_hw2_yw4251"
author: "You Wu"
date: 2023-10-01
output: github_document
---
# Problem 0
```{r}
library(tidyverse)
library(readxl)
```


# Problem 1
Clean `pols-month.csv`. Apply function`month.name`to replace month number with month name.
```{r}
pols_month_df=
  read_csv("data/pols-month.csv")|>
  separate(mon,into=c("year","month","day"),convert=TRUE)|>
  mutate(
    month=month.name[as.numeric(month)],
    president=recode(prez_gop,"0" = "dem", "1" = "gop", "2" = "gop")
  )|>
  select(-prez_dem,-prez_gop,-day)
  
```
Similarly, clean `snp.csv`.
```{r}
snp_df=
  read_csv("data/snp.csv")|>
  mutate(date = as.Date(date, format = "%m/%d/%y"))|>
  separate(date,into=c("year","month","day"),convert=TRUE)|>
  arrange(year,month)|>
  mutate(
    month=month.name[as.numeric(month)]
  )|>
  select(year,month,close)
```
Next, Clean `Unemployment data`.

```{r}
unemp_df=
  read_csv('data/unemployment.csv')|>
  pivot_longer(
    Jan:Dec,
    names_to="month",
    values_to="percentage"
  )|>
  rename(year=Year)
  
```
Now merge:

```{r}
df=left_join(pols_month_df,snp_df)|>
  mutate(month=substr(month,1,3))|>
  left_join(x= _,unemp_df)

str(df)
```

## Discussion:

The `pols-month` file comprises `r nrow(pols_month_df)` observations across `r ncol(pols_month_df)` variables, providing insights into the count of national politicians affiliated with the Democratic or Republican parties from  years `r range(pull(pols_month_df, year))[1]` to `r range(pull(pols_month_df, year))[2]`. The `snp` file encompasses `r nrow(snp_df)` observations, featuring `r ncol(snp_df)` variables that offer insights into Standard & Poor's stock market index (S&P) ranging from years `r range(pull(snp_df, year))[1]` to `r range(pull(snp_df, year))[2]`. Widely recognized as a representative measure of the overall stock market, these variables provide valuable information on market trends and performance. The file `unemployment` contains `r nrow(unemp_df)` observations of `r ncol(unemp_df)` variables ranging from years `r range(pull(unemp_df, year))[1]` to `r range(pull(unemp_df, year))[2]`. The average unemployment rate is `r mean(na.omit(pull(unemp_df, percentage)))`%.

For the resulting `df`, it has `r nrow(df)` observations across `r ncol(df)` variables with years ranging from `r range(pull(df, year))[1]` to `r range(pull(df, year))[2]`. Variables include:

**_gov_gop:_** the number of republican governors on the associated date

**_sen_gop:_** the number of republican senators on the associated date

**_rep_gop:_** the number of republican representatives on the associated date

**_gov_dem:_** the number of democratic governors on the associated date

**_sen_dem:_** the number of democratic senators on the associated date

**_rep_dem:_** the number of democratic representatives on the associated date

**_president:_**: the party affiliation of the politician

**_close:_** the closing values of the S&P stock index on the associated date. The average closing rate is `r round(mean(na.omit(pull(df, close))),2)`.

**_percentage:_** the percentage of unemployment.The average unemployment rate is `r mean(na.omit(pull(df, percentage)))`.

# Problem 2
Clean `Mr. Trash Wheel`.
```{r}
mr_df=
  readxl::read_excel("data/202309 Trash Wheel Collection Data.xlsx",sheet=1,range="A2:N587")|>
  janitor::clean_names()|>
  filter(!is.na(dumpster))|>
  mutate(homes_powered=weight_tons*500/30,
         trash_wheel="mr",
         year=as.numeric(year))
```

Clean `Professor Trash Wheel`.
```{r}
prof_df=
  readxl::read_excel("data/202309 Trash Wheel Collection Data.xlsx",sheet=2,range="A2:M109")|>
  janitor::clean_names()|>
  filter(!is.na(dumpster))|>
  mutate(homes_powered=weight_tons*500/30,
         trash_wheel="professor")
```

Clean `Gwynnda Trash Wheel`
```{r}
gwy_df=
  readxl::read_excel("data/202309 Trash Wheel Collection Data.xlsx",sheet=4,range="A2:L159")|>
  janitor::clean_names()|>
  filter(!is.na(dumpster))|>
  mutate(homes_powered=weight_tons*500/30,
         trash_wheel="gwynnda")
  
```


Then combine these datasets.

```{r}
combined_df=bind_rows(mr_df,prof_df,gwy_df)|>
  mutate(homes_powered=round(homes_powered,digits=0))
str(combined_df)
```

## Description:

The `mr_df`comprises `r nrow(mr_df)` observations across `r ncol(mr_df)` variables. Key variables include `dumpster`, `date`, `weight tons`,`homes powered` and `trash wheel (name)`.

The `prof_df`comprises `r nrow(prof_df)` observations across `r ncol(prof_df)` variables. Key variables include `dumpster`, `date`, `weight tons`,`homes powered` and `trash wheel (name)`.

The `gwy_df`comprises `r nrow(gwy_df)` observations across `r ncol(gwy_df)` variables. Key variables include `dumpster`, `date`, `weight tons`,`homes powered` and `trash wheel (name)`.

The `combined_df` comprises `r nrow(combined_df)` observations across `r ncol(combined_df)` variables, ranging from years `r range(combined_df$year)[1]` to `r range(combined_df$year)[2]`. The combined dataset includes information on the `dumpster number`, `date of collection`, `amount of total litter` and `litter type` as well as `three different trash wheel(Mr. Trash Wheel, Professor Trash wheel, Gwynnda Trash Wheel)`. The total weight of trash collected by Professor Trash Wheel is `r filter(combined_df, trash_wheel=='professor')|>pull(weight_tons)|>sum()` tons. The total number of cigarette butts collected by Gwynnda in July of 2021 is `r filter(combined_df, trash_wheel=="gwynnda",year==2021,month=="July")|>pull(cigarette_butts)|>sum()`.

# Problem 3

Import `MCI baseline` data, and skip the first row of description. I use `janitor` function to standardize column names. Then convert `sex` variable and `apoe4` variable to factor variable, and convert `age at onset` variable to numeric variable. Then use `filter` function to drop participants who do not meet the stated inclusion criteria.
```{r}
baseline_df=
  read_csv("data/MCI_baseline.csv",skip=1)|>
  janitor::clean_names()|>
  mutate(
    sex=as.factor(sex),
    apoe4=as.factor(apoe4),
    age_at_onset=as.numeric(age_at_onset)
  )|>
  filter(!is.na(age_at_onset))
```

## Discussion

The `MCI baseline` data comprises `r nrow(baseline_df)` observations across `r ncol(baseline_df)` variables. The dataset contains information of `current age`, `sex`, `years of education`, `whether APOE4 carrier` and `age at onset of MCI`. For participants who develop MCI, their average baseline age is `r round(mean(baseline_df$current_age),2)`, and their current age ranges from `r range(baseline_df$current_age)[1]` to `r range(baseline_df$current_age)[2]`. Altogether, `r nrow(read_csv("data/MCI_baseline.csv",skip=1))` participates are recruited. Among them, `r nrow(baseline_df)` develop MCI. In the study, `r round(nrow(filter(baseline_df, sex==0,apoe4==1))/nrow(filter(baseline_df,sex==0))*100,2)`%  women are APOE4 carriers.

Next, clean `MCI amyloid` data, and skip the first row of description. I use `janitor` function to standardize column names. Then, I rename the `study_id` variable to `id` for merging purpose and `baseline` variable to `time_0` for consistency. `Pivot longer` function was used to tidy the dataset of longitudinally observed biomarker values. Finally, turn all columns to numeric variable.


```{r}
amyloid_df=
  read_csv("data/mci_amyloid.csv",skip=1,na=c('Na',"NA"))|>
  janitor::clean_names()|>
  rename(id=study_id,time_0=baseline)|>
  pivot_longer(
    time_0:time_8,
    names_to="time",
    values_to="42/40 ratio",
    names_prefix = ("time_")
  )|>
  as.data.frame(lapply(amyloid_df, as.numeric))
```

## Discussion

The `amyloid_df` comprises `r nrow(amyloid_df)` observations across `r ncol(amyloid_df)` variables. The dataset keeps record of biomarker Amyloid β42/40 ratio in different time. Key variables are `id`, `time` and `β42/40 ratio`. The `β42/40 ratio` variable has some `NA` values will be dropped in the next step.

Now check whether some participants appear in only the baseline or amyloid datasets.

```{r}
only_baseline=setdiff(baseline_df$id,amyloid_df$id)
only_amyloid=setdiff(amyloid_df$id,baseline_df$id)
```

As shown, id of `r only_baseline` only appear in the baseline dataset. Id of `r only_amyloid` only appear in the amyloid dataset, altogether `r length(only_amyloid)` ids.

Time to merge these two datasets! Rows with `NA` are dropped.

```{r}
merged_df=merge(baseline_df,amyloid_df,by="id",all=FALSE)|>drop_na()
write_csv(merged_df,"data/merged_df.csv")
str(merged_df)
```

The `merged_df` comprises `r nrow(merged_df)` observations across `r ncol(merged_df)` variables. Key variables include `id`, `current age`, `sex`, `education of year`, `whether apoe4 carrier`, `age at onset`, `time` and `amyloid ratio`.




